From Precipitation to Prediction: Using ConvLSTM Models for Comprehensive River Runoff Forecasting

### 2024-06-07 paper/paper.tex
Update

\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={From Precipitation to Prediction: Integrating ConvLSTM Models for Comprehensive River Runoff Forecasting},
  pdfauthor={Florian Börgel; Sven Karsten; Karoline Rummel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\title{From Precipitation to Prediction: Integrating ConvLSTM Models for
Comprehensive River Runoff Forecasting}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi

@@

River runoff is an important component of the global water cycle as it
comprises about one third of the precipitation over land areas (Hagemann
et al., 2020). In the context of climate change studies, river runoff is
usually generated in two ways. First, river runoff as input for ocean
models can be created using hydrological models such as the Hydrological
Discharge (HD) model (Hagemann et al., 2020). HD models calculates the
water balance using hydrological processes (e.g.~snow, glaciers, soil
moisture, groundwater contribution). It represents a complex forecasting
tool that uses underlying physical processes. A different approach would
use data-based models that integrate statistical correction, using the
land surface schemes of global or regional climate models.

The relatively recent emergence of machine learning (ML) models has
offered different approaches to river runoff forecasting. Accurate
runoff forecasting is crucial for effective water resources management,
particularly over extended periods (W. Fang et al., 2019; Tan et al.,
2018). Common approaches employ feed-forward artificial neural networks,
support vector machines, adaptive neuro-fuzzy inference systems, and
notably, Long Short-Term Memory (LSTM) neural networks that have gained
traction for long-term hydrological forecasting due to their excellent
performance (Ashrafi et al., 2017; L. Fang \& Shao, 2022; Huang et al.,
2014; Humphrey et al., 2016; Kratzert et al., 2018; Liu et al., 2020).

LSTM networks, first introduced by Hochreiter \& Schmidhuber (1997) ,
are an evolution of the classical Recurrent Neural Networks (RNNs).
Their structure enables them to learn long-term dependencies while
avoiding the vanishing or exploding gradient problem. They have shown
stability and efficacy in sequence-to-sequence predictions. However, a
limitation of LSTMs is their inability to effectively capture
two-dimensional structures, an area where Convolutional Neural Networks
(CNNs) excel. Recognizing this limitation SHI et al. (2015) proposed a
Convolutional LSTM (ConvLSTM) architecture, which combines the strengths
of both LSTM and CNN. The ConvLSTM network has been proven useful for
spatio-temporal applications such as precipitation nowcasting (SHI et
al., 2015), flood forecasting (Moishin et al., 2021), and river runoff
forecasting (Ha et al., 2021; Zhu et al., 2023).

In this work, we demonstrate that ConvLSTM networks are a reliable
method for predicting multiple rivers simultaneously, using only
atmospheric forcing, even in the absence of a fully functional
hydrological model with a complex parameterization. We use the Baltic
Sea catchment as an example to illustrate our approach. Although the
methodology we propose is universally applicable across various
geographic regions, the Baltic Sea presents a challenge due to its
unique hydrological characteristics, being nearly decoupled from the
open ocean (see Figure). As a consequence, the salinity of the Baltic
Sea is driven to a large part by freshwater supply from rivers.
Freshwater enters the Baltic Sea through river runoff or positive net
precipitation (precipitation minus evaporation) over the sea surface.
The net precipitation accounts for 11 \(\%\) and the river input for 89
\(\%\) of the total freshwater input (Meier \& Döscher, 2002). Modelling
the Baltic Sea is therefore to a large part the result of the quality of
the river input, that is used for the simulation. This makes the
accurate modeling of river runoff especially critical for simulations
pertaining to the Baltic Sea.

@@

As an alternative, we propose that this functional can be adequately
represented by a combination of a ConvLSTM with a subsequent FC network.
This approach eliminates the need for detailed knowledge of the involved
processes and their modeling. Instead, these features can be ``learned''
by the network in an automated manner. Our proposed network architecture
is visualized in Figure~\ref{fig-ConvLSTM_FC} and described in detail in
the following sections. To provide an overview, we will discuss the main
components of this architecture one-by-one.

@@

Before turning directly to the convolutional LSTM, the simpler
architecture of the plain LSTM is examined which serves as a foundation
for understanding the more complex ConvLSTM.

The Long Short-Term Memory (LSTM), a specialized form of Recurrent
Neural Networks (RNNs), is specifically designed to model temporal
sequences
\(\vec{X}^t[1], \ldots \vec{X}^t[\tau],\ldots \vec{X}^t[N_\tau]\) of
\(N_\tau\) input quantities
\(\vec{X}^t[\tau] = (X^t_k[\tau]) \in \mathbb{R}^{N_k}\). This sequence
is taken from a dataset given in form of a time series \(\{\vec{X}^t\}\)
with the endpoint coinciding with the specific element in the time
series connected to time \(t\),
i.e.~\(\vec{X}^t[N_\tau] \equiv \vec{X}^t\), see
Figure~\ref{fig-ConvLSTM_FC}. Here, \(N_k\) represents the number of
input ``channels,'' which can correspond to different measurable
quantities. The LSTM's unique design allows it to adeptly handle
long-range dependencies, setting it apart from traditional RNNs in terms
of accuracy (see Figure~\ref{fig-lstm}).

@@

Second, the actual gate vectors are computed by the core equations of
the LSTM as proposed by Hochreiter \& Schmidhuber (1997):

@@

The runoff data covering the period 1979 to 2011 is based on an E-HYPE
hindcast simulation that was forced by a regional downscaling of
ERA-Interim (Dee et al., 2011) with RCA3 ({``The rossby centre regional
climate model RCA3,''} n.d.) and implemented into NEMO-Nordic (Hordoir
et al., 2019) as a mass flux. For the periods before (1961 to 1978) and
after (2012 to 2018) additional spatial temporal corrections have been
applied to the runoff data, and have therefore been ignored. The quality
of the runoff was extensively evaluated. For more information see Gröger
et al. (2022) and references therein.

@@

To simulate the Baltic Sea, we use a coupled 3-dimensional ocean model
Baltic Sea, called the Modular Ocean Model (MOM). This model uses a
finite-difference method to solve the full set of primitive equations to
calculate the motion of water and the transport of heat and salt. The
K-profile parameterization (KPP) was used as turbulence closure scheme.
The model's western boundary opens into the Skagerrak and connects the
Baltic Sea to the North Sea. The maximum depth was set at 264 meters. A
more detailed description of the setup can be found in (Gröger et al.,
2022).

@@

As input the model receives 30 days of atmospheric surface fields
temperature \(T\), precipitation \(P\), specific humidity \(Q\) and wind
speed \(W\), with a daily resolution to predict the river runoff \(R\)
at the time step \(\Delta t+1\) , which can be summarized as

\(R_{\Delta t+1} = f\left(T_{t-30:t}, P_{t-30:t}, Q_{t-30:t}, W_{t-30:t}\right)\)

with f being a function maps the 30 days of daily atmospheric surface
fields data to the predicted river runoff.

The choice of atmospheric fields was based on the implemented river
runoff calculation in the atmospheric model COSMO-CLM which uses these
atmospheric fields to calculate an river runoff estimate.

@@

The accuracy of the model is displayed in Figure
\textbf{?@fig-statistical-evaluationNN}. As mention above for
evaluation, the test dataset was utilized. The left panel panel (Figure
\textbf{?@fig-statistical-evaluationNN} a) ) illustrates the relative
prediction error in relation to the original E-HYPE data, indicating
that, on daily timescales, the model can predict river runoff with an
accuracy of \(\pm\) 5\(\%\). The overall correlation is 0.997 with the
resulting error metrics yielding a RMSE of 323.99 \(m^3\)/s and MAE of
249.51 \(m^3\)/s. While the model's performance is satisfactory, the
discrepancies between the actual values and the predictions could partly
be attributed to the use of a different atmospheric dataset than the one
originally used to drive the E-HYPE model. However, by applying a
rolling mean with a 5-day window, the prediction error is reduced to
less than 1\(\%\), which is acceptable for the purposes of climate
modeling. Lastly, the right panel demonstrates that the distribution of
residuals follows a Gaussian shape, suggesting that our model does not
exhibit bias by systematically over or underestimating river runoff
values.

@@

Phasellus interdum tincidunt ex, a euismod massa pulvinar at. Ut
fringilla ut nisi nec volutpat. Morbi imperdiet congue tincidunt.
Vivamus eget rutrum purus. Etiam et pretium justo. Donec et egestas sem.
Donec molestie ex sit amet viverra egestas. Nullam justo nulla,
fringilla at iaculis in, posuere non mauris. Ut eget imperdiet elit.

\hypertarget{open-research}{%
\section{Open research}\label{open-research}}

Phasellus interdum tincidunt ex, a euismod massa pulvinar at. Ut
fringilla ut nisi nec volutpat. Morbi imperdiet congue tincidunt.
Vivamus eget rutrum purus. Etiam et pretium justo. Donec et egestas sem.
Donec molestie ex sit amet viverra egestas. Nullam justo nulla,
fringilla at iaculis in, posuere non mauris. Ut eget imperdiet elit.

@@

\leavevmode\vadjust pre{\hypertarget{ref-ashrafi2017fully}{}}%
Ashrafi, M., Chua, L. H. C., Quek, C., \& Qin, X. (2017). A fully-online
neuro-fuzzy model for flow forecasting in basins with limited data.
\emph{Journal of Hydrology}, \emph{545}, 424--435.

\leavevmode\vadjust pre{\hypertarget{ref-dee2011}{}}%
Dee, D. P., Uppala, S. M., Simmons, A. J., Berrisford, P., Poli, P.,
Kobayashi, S., et al. (2011). The ERA-Interim reanalysis: configuration
and performance of the data assimilation system. \emph{Quarterly Journal
of the Royal Meteorological Society}, \emph{137}(656), 553--597.
\url{https://doi.org/10.1002/qj.828}

\leavevmode\vadjust pre{\hypertarget{ref-fang2022application}{}}%
Fang, L., \& Shao, D. (2022). Application of long short-term memory
(LSTM) on the prediction of rainfall-runoff in karst area.
\emph{Frontiers in Physics}, \emph{9}, 790687.

@@

\leavevmode\vadjust pre{\hypertarget{ref-gruxf6ger2022}{}}%
Gröger, M., Placke, M., Meier, M., Börgel, F., Brunnabend, S.-E.,
Dutheil, C., et al. (2022). \emph{The Baltic Sea model inter-comparison
project BMIP {\textendash} a platform for model development, evaluation,
and uncertainty assessment}. Retrieved from
\url{https://gmd.copernicus.org/preprints/gmd-2022-160/}

\leavevmode\vadjust pre{\hypertarget{ref-ha2021}{}}%
Ha, S., Liu, D., \& Mu, L. (2021). Prediction of Yangtze River
streamflow based on deep learning neural network with El
Niño{\textendash}Southern Oscillation. \emph{Scientific Reports},
\emph{11}(1), 11738. \url{https://doi.org/10.1038/s41598-021-90964-3}

@@

\leavevmode\vadjust pre{\hypertarget{ref-hochreiter1997long}{}}%
Hochreiter, S., \& Schmidhuber, J. (1997). Long short-term memory.
\emph{Neural Computation}, \emph{9}(8), 1735--1780.

\leavevmode\vadjust pre{\hypertarget{ref-hordoir2019}{}}%
Hordoir, R., Axell, L., Höglund, A., Dieterich, C., Fransner, F.,
Gröger, M., et al. (2019). Nemo-nordic 1.0: A NEMO-based ocean model for
the baltic and north seas {\textendash} research and operational
applications. \emph{Geoscientific Model Development}, \emph{12}(1),
363--386. \url{https://doi.org/10.5194/gmd-12-363-2019}

\leavevmode\vadjust pre{\hypertarget{ref-huang2014monthly}{}}%
Huang, S., Chang, J., Huang, Q., \& Chen, Y. (2014). Monthly streamflow
prediction using modified EMD-based support vector machine.
\emph{Journal of Hydrology}, \emph{511}, 764--775.

\leavevmode\vadjust pre{\hypertarget{ref-humphrey2016hybrid}{}}%
Humphrey, G. B., Gibbs, M. S., Dandy, G. C., \& Maier, H. R. (2016). A
hybrid approach to monthly streamflow forecasting: Integrating
hydrological model outputs into a bayesian artificial neural network.
\emph{Journal of Hydrology}, \emph{540}, 623--640.

\leavevmode\vadjust pre{\hypertarget{ref-kratzert2018rainfall}{}}%
Kratzert, F., Klotz, D., Brenner, C., Schulz, K., \& Herrnegger, M.
(2018). Rainfall--runoff modelling using long short-term memory (LSTM)
networks. \emph{Hydrology and Earth System Sciences}, \emph{22}(11),
6005--6022.

\leavevmode\vadjust pre{\hypertarget{ref-liu2020streamflow}{}}%
Liu, D., Jiang, W., Mu, L., \& Wang, S. (2020). Streamflow prediction
using deep learning neural network: Case study of yangtze river.
\emph{IEEE Access}, \emph{8}, 90069--90086.

\leavevmode\vadjust pre{\hypertarget{ref-meier2002simulated}{}}%
Meier, H. M., \& Döscher, R. (2002). Simulated water and heat cycles of
the baltic sea using a 3D coupled atmosphere--ice--ocean model.
\emph{Boreal Environment Research}, \emph{7}(4), 327.

\leavevmode\vadjust pre{\hypertarget{ref-moishin2021designing}{}}%
Moishin, M., Deo, R. C., Prasad, R., Raj, N., \& Abdulla, S. (2021).
Designing deep-based learning flood forecast model with ConvLSTM hybrid
algorithm. \emph{IEEE Access}, \emph{9}, 50982--50993.

@@

\leavevmode\vadjust pre{\hypertarget{ref-theross}{}}%
The rossby centre regional climate model RCA3: Model description and
performance - tellus a: Dynamic meteorology and oceanography. (n.d.).
Retrieved from
\url{https://a.tellusjournals.se/articles/10.1111/j.1600-0870.2010.00478.x}

\leavevmode\vadjust pre{\hypertarget{ref-zhu2023spatiotemporal}{}}%
Zhu, S., Wei, J., Zhang, H., Xu, Y., \& Qin, H. (2023). Spatiotemporal
deep learning rainfall-runoff forecasting combined with remote sensing
precipitation products in large scale basins. \emph{Journal of
Hydrology}, \emph{616}, 128727.

>>>>

\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={From Precipitation to Prediction: Using ConvLSTM Models for Comprehensive River Runoff Forecasting},
  pdfauthor={Florian Börgel; Sven Karsten; Karoline Rummel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\title{From Precipitation to Prediction: Using ConvLSTM Models for
Comprehensive River Runoff Forecasting}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, enhanced, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

River runoff is an important component of the global water cycle as it
comprises about one third of the precipitation over land areas (Hagemann
et al., 2020). Therefore, accurate runoff forecasting is crucial for
effective water resources management, particularly over extended periods
(Fang et al., 2019; Tan et al., 2018). In the context of climate change
studies, river runoff is usually generated in two ways. First, river
runoff as input for ocean models can be created using hydrological
models such as the Hydrological Discharge (HD) model (Hagemann et al.,
2020). HD models calculate the water balance using hydrological
processes (e.g.~snow, glaciers, soil moisture, groundwater
contribution). It represents a complex forecasting tool that uses
underlying physical processes. The second approach uses data-driven
models that integrate statistical correction, using land surface schemes
of global or regional climate models {[}QUELLE{]}.

With the recent rise of machine learning in climate research various
model architectures have also been tested for river runoff forecasting.
Common approaches employ feed-forward artificial neural networks,
support vector machines, adaptive neuro-fuzzy inference systems, and
notably, Long Short-Term Memory (LSTM) neural networks that have gained
traction for long-term hydrological forecasting due to their excellent
performance (\textbf{humphrey2016hybrid?}; \textbf{huang2014monthly?};
\textbf{ashrafi2017fully?}; \textbf{liu2020streamflow?};
\textbf{fang2022application?}; \textbf{kratzert2018rainfall?}).

LSTM networks, first introduced by (\textbf{hochreiter1997long?}) , are
an evolution of the classical Recurrent Neural Networks (RNNs). Their
structure enables them to learn long-term dependencies while avoiding
the vanishing or exploding gradient problem. They have shown stability
and efficacy in sequence-to-sequence predictions. However, a limitation
of LSTMs is their inability to effectively capture two-dimensional
structures, an area where Convolutional Neural Networks (CNNs) excel.
Recognizing this limitation SHI et al. (2015) proposed a Convolutional
LSTM (ConvLSTM) architecture, which combines the strengths of both LSTM
and CNN. The ConvLSTM network has been proven useful for spatio-temporal
applications such as precipitation nowcasting (SHI et al., 2015), flood
forecasting (\textbf{moishin2021designing?}), and river runoff
forecasting (\textbf{ha2021?}; \textbf{zhu2023spatiotemporal?}).

In this work, we demonstrate that ConvLSTM networks are a reliable
method for predicting multiple rivers simultaneously, using only
atmospheric forcing, even in the absence of a fully functional
hydrological model with a complex parameterization. We use the Baltic
Sea catchment as an example to illustrate our approach. Although the
methodology we propose is universally applicable across various
geographic regions, the Baltic Sea represents a challenging region due
to its unique hydrological characteristics, being nearly decoupled from
the open ocean (see Figure). As a consequence, the salinity of the
Baltic Sea is driven to a large part by freshwater supply from rivers.
Freshwater enters the Baltic Sea through river runoff or positive net
precipitation (precipitation minus evaporation) over the sea surface.
The net precipitation accounts for 11 \(\%\) and the river input for 89
\(\%\) of the total freshwater input (\textbf{meier2002simulated?}).
Modelling the Baltic Sea is therefore to a large part the result of the
quality of the river input, that is used for the simulation. This makes
the accurate modeling of river runoff especially critical for
simulations pertaining to the Baltic Sea.

@@

As an alternative, we propose that this functional can be adequately
represented by a combination of a convolutional Long-Short Term Memory
(ConvLSTM) model with a subsequent fully connected neural network. This
approach eliminates the need for detailed knowledge of the involved
processes and their modeling. Instead, these features can be ``learned''
by the network in an automated manner. Our proposed network architecture
is visualized in Figure~\ref{fig-ConvLSTM_FC} and described in detail in
the following sections. To provide an overview, we will discuss the main
components of this architecture one-by-one.

@@

Before turning directly to the ConvLSTM the simpler architecture of the
plain Long-Short Term Memory (LSTM) model is examined which serves as a
foundation for understanding the more complex ConvLSTM.

The LSTM, a specialized form of Recurrent Neural Networks (RNNs), is
specifically designed to model temporal sequences
\(\vec{X}^t[1], \ldots \vec{X}^t[\tau],\ldots \vec{X}^t[N_\tau]\) of
\(N_\tau\) input quantities
\(\vec{X}^t[\tau] = (X^t_k[\tau]) \in \mathbb{R}^{N_k}\). This sequence
is taken from a dataset given in form of a time series \(\{\vec{X}^t\}\)
with the endpoint coinciding with the specific element in the time
series connected to time \(t\),
i.e.~\(\vec{X}^t[N_\tau] \equiv \vec{X}^t\), see
Figure~\ref{fig-ConvLSTM_FC}. Here, \(N_k\) represents the number of
input ``channels,'' which can correspond to different measurable
quantities. The LSTM's unique design allows it to adeptly handle
long-range dependencies, setting it apart from traditional RNNs in terms
of accuracy (see Figure~\ref{fig-lstm}).

@@

Second, the actual gate vectors are computed by the core equations of
the LSTM as proposed by (\textbf{hochreiter1997long?}):

@@

The runoff data covering the period 1979 to 2011 is based on an E-HYPE
hindcast simulation that was forced by a regional downscaling of
ERA-Interim (\textbf{dee2011?}) with RCA3 (\textbf{theross?}) and
implemented into NEMO-Nordic (\textbf{hordoir2019?}) as a mass flux. For
the periods before (1961 to 1978) and after (2012 to 2018) additional
spatial temporal corrections have been applied to the runoff data, and
have therefore been ignored. The quality of the runoff was extensively
evaluated. For more information see (\textbf{gröger2022?}) and
references therein.

@@

To simulate the Baltic Sea, we use a coupled 3-dimensional ocean model
Baltic Sea, called the Modular Ocean Model (MOM). This model uses a
finite-difference method to solve the full set of primitive equations to
calculate the motion of water and the transport of heat and salt. The
K-profile parameterization (KPP) was used as turbulence closure scheme.
The model's western boundary opens into the Skagerrak and connects the
Baltic Sea to the North Sea. The maximum depth was set at 264 meters. A
more detailed description of the setup can be found in
(\textbf{gröger2022?}).

@@

As input the model receives \(N_{\tau}\) = 30 days of atmospheric
surface fields temperature \(T\), precipitation \(P\), specific humidity
\(Q\) and wind speed \(W\), with a daily resolution to predict the river
runoff \(\vec{R}\). The choice of atmospheric fields was based on the
implemented river runoff calculation in the atmospheric model COSMO-CLM
which uses these atmospheric fields to calculate an river runoff
estimate.

@@

The accuracy of the model is displayed in Figure
\textbf{?@fig-statistical-evaluationNN}. As mention above for
evaluation, the test dataset was utilized. The left panel panel (Figure
\textbf{?@fig-statistical-evaluationNN} a) illustrates the relative
prediction error in relation to the original E-HYPE data, indicating
that, on daily timescales, the model can predict river runoff with an
accuracy of \(\pm\) 5\(\%\). The overall correlation is 0.997 with the
resulting error metrics yielding a RMSE of 323.99 \(m^3\)/s and MAE of
249.51 \(m^3\)/s. While the model's performance is satisfactory, the
discrepancies between the actual values and the predictions could partly
be attributed to the use of a different atmospheric dataset than the one
originally used to drive the E-HYPE model. However, by applying a
rolling mean with a 5-day window, the prediction error is reduced to
less than 1\(\%\), which is acceptable for the purposes of climate
modeling. Lastly, the right panel demonstrates that the distribution of
residuals follows a Gaussian shape, suggesting that our model does not
exhibit bias by systematically over or underestimating river runoff
values.

@@

The research presented in this study is part of the Baltic Earth program
(Earth System Science for the Baltic Sea region, see
\href{https://www.baltic.earth/}{https://www.baltic.earth}.

@@

@@

@@

@@
